---
title: "fos-net network visualization"
author: "Maureen Ritchey"
output:
  html_notebook:
    theme: cosmo
    toc: yes
    toc_float: yes
  html_document:
    theme: cosmo
    toc: yes
    toc_float: yes
---


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(network)
library(sna)
library(ggnet)
```

# Load & tidy data

All of the data were generated using the Brain Connectivity Toolbox in Matlab. I'm importing it into R because it will make it easier to make the plots and run stats.

```{r}
rawdata <- read.csv('data/Fos counts - clean.csv')
fosnet <- read.csv('data/fosnet_matrix_ALL.csv')
nodevals <- read.csv('data/fosnet_nodevals_ALL.csv')

fosnet$condition <- 'ALL'
nodevals$condition <- 'ALL'
targ.mod <- fosnet$target.module
seed.mod <- fosnet$seed.module

# Get the condition-specific data and add it to the dfs
gatherData <- function(fosnet,whichdf,label) {
  fosnet.new <- read.csv(paste0('data/fosnet_',whichdf,'_',label,'.csv'))
  fosnet.new$condition <- label
  if (whichdf=="matrix") {
  fosnet.new$target.module <- targ.mod # replacing this because modules defined only by ALL
  fosnet.new$seed.module <- seed.mod
  }
  fosnet <- rbind(fosnet,fosnet.new)
}
fosnet <- gatherData(fosnet,'matrix','SJ')
fosnet <- gatherData(fosnet,'matrix','SA')
fosnet <- gatherData(fosnet,'matrix','NJ')
fosnet <- gatherData(fosnet,'matrix','Nad')

nodevals <- gatherData(nodevals,'nodevals','SJ')
nodevals <- gatherData(nodevals,'nodevals','SA')
nodevals <- gatherData(nodevals,'nodevals','NJ')
nodevals <- gatherData(nodevals,'nodevals','Nad')

rawdata.long <- gather(rawdata,key=region,value=fos,10:38)

# Add label info to dataframes
seedkey <- 
  select(fosnet,seed,seed.module,seed.group) %>%
  mutate(seed.label = factor(paste0(seed.module,'.',seed.group,'.',seed)),
         seed.ins = ifelse(seed=="AI"|seed=="DI"|seed=="GI","Insula","Not insula")) %>%
  unique() %>%
  arrange(seed.module,seed.group,seed) 

targetkey <- 
  mutate(seedkey,
         target.module = seed.module,
         target.label = seed.label,
         target = seed,
         target.group = seed.group,
         target.ins = seed.ins) %>%
  select(target,target.label,target.group,target.ins,target.module)

fosnet %>%
  left_join(seedkey) %>%
  left_join(targetkey) -> fosnet

fosnet$condition[fosnet$condition=="Nad"] <- "NA"
fosnet$condition <- factor(fosnet$condition,levels=c("ALL","NJ","NA","SJ","SA"))

nodevals$condition[nodevals$condition=="Nad"] <- "NA"
nodevals$condition <- factor(nodevals$condition,levels=c("ALL","NJ","NA","SJ","SA"))

```

# Plot network, sorted by module assignment
## Plot connection matrix

This is the "connectivity" matrix, sorted by module assignment (as determined by a BCT community detection algorithm).

```{r, fig.height=3, fig.width=3.5}

# Getting ordered labels from ALL
tmp <- fosnet[fosnet$condition=="ALL",]$target.label 
myOrders.targ <- as.numeric(droplevels(tmp))

tmp <- fosnet[fosnet$condition=="ALL",]$seed.label 
myOrders.seed <- as.numeric(droplevels(tmp))

# Function for plotting correlation matrix
plotCorrMat <- function(fosnet,cond,lims=c(-1,1)) {
  
  fosnet %>%
    filter(condition==cond) %>%
    mutate(target.ordered=reorder(target,myOrders.targ), # this puts the targets in the same order
           seed.ordered=reorder(seed,myOrders.seed),
           seed.ordered=factor(seed.ordered,levels=rev(levels(seed.ordered))),
           correlation=connection) %>% # this reverses seed so it goes from left top to bottom right
    ggplot(aes(x=target.ordered,y=seed.ordered)) + geom_tile(aes(fill=correlation),colour = "white") + 
    scale_fill_gradient2(low = "blue",mid="white",high = "red",midpoint=0,limits=lims) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, color="black")) +
    theme(axis.text.y = element_text(color="black")) +
    xlab("") + ylab("") + ggtitle(cond) -> p
  
  print(p)
  
  ggsave(paste0('figs/correlation_matrix_',cond,'.eps'),device="eps")
  
}


plotCorrMat(fosnet,'ALL',c(-.7,.7))
plotCorrMat(fosnet,'NJ')
plotCorrMat(fosnet,'NA')
plotCorrMat(fosnet,'SJ')
plotCorrMat(fosnet,'SA')

# Saving out the level orders for later...
fosnet %>%
    filter(condition=="ALL") %>%
    mutate(target.ordered=reorder(target,myOrders.targ), # this puts the targets in the same order
           seed.ordered=reorder(seed,myOrders.seed),
           seed.ordered=factor(seed.ordered,levels=rev(levels(seed.ordered))),
           correlation=connection) -> tmp
tmplevels <- levels(tmp$seed.ordered)


```

### Summarize within and between module correlations
```{r, fig.height=2.5, fig.width=1}
fosnet %>%
  mutate(conntype = factor(ifelse(seed.module==1 & target.module==1,'within module 1',
                                  ifelse(seed.module==2 & target.module==2,'within module 2','between modules')),
                           levels=c("within module 1","within module 2","between modules")),
         connection = ifelse(seed==target,NA,connection)) %>%
  filter(!is.na(connection)) %>%
  group_by(condition,conntype) %>%
  summarize(conn = mean(connection,na.rm=TRUE),
            conn_n = length(!is.na(connection)),
            connsem = sd(connection,na.rm=TRUE)/sqrt(length(!is.na(connection))/2), # dividing by two bc symmetric matrix
            connsd = sd(connection,na.rm=TRUE),
            connmin = min(connection,na.rm=TRUE),
            connmax = max(connection,na.rm=TRUE)) %>%  
  arrange(conntype,condition) -> summarymodcorr

fosnet %>%
  mutate(conntype = factor(ifelse(seed.module==1 & target.module==1,'within module 1',
                                  ifelse(seed.module==2 & target.module==2,'within module 2','between modules')),
                           levels=c("within module 1","within module 2","between modules")),
         conn = ifelse(seed==target,NA,connection)) %>%
  filter(!is.na(connection)) -> mod.replicates

print(summarymodcorr)

summarymodcorr %>%
  ggplot(aes(x=condition,y=conn,fill=condition)) + geom_bar(stat="identity") + facet_wrap(~conntype,nrow=3,ncol=1) +
  geom_errorbar(aes(x=condition,ymin=conn-connsem,ymax=conn+connsem),width=.25) + guides(fill=FALSE) + xlab('')

ggsave(filename='figs/with_betw_mod_orig.png',device='png')


mod.replicates %>%
  ggplot(aes(x=condition,y=conn,fill=condition)) + geom_jitter(size=.4,alpha=.4,width=.1,aes(color=condition)) + 
  facet_wrap(~conntype,nrow=3,ncol=1) +
  guides(fill=FALSE,color=FALSE) + xlab('') +
  geom_crossbar(mapping = aes(x = condition, y = conn),
                  stat = "summary",
                  fun.ymin = function(z) {quantile(z,0.25)},
                  fun.ymax = function(z) {quantile(z,0.75)},
                  fun.y = median,
                  size=.3, alpha=0, width=.2) + theme_minimal() +
  ylab("correlation")

ggsave(filename='figs/with_betw_mod_quant.png',device='png',dpi=1200)


```

```{r, fig.height=1, fig.width=3.5}
fosnet %>%
  group_by(condition,seed.module,target.module) %>%
  summarize(conn = mean(connection,na.rm=TRUE)) %>%
  mutate(seed.mod=factor(seed.module,levels=c("2","1")),
         target.mod=factor(target.module)) %>%
  ggplot(aes(x=target.mod,y=seed.mod)) + 
  geom_tile(aes(fill=conn),colour = "white") + 
  scale_fill_gradient2(low = "blue",mid="white",high = "red",midpoint=0) +
  facet_grid(.~condition) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, color="black")) +
  theme(axis.text.y = element_text(color="black")) + xlab('') + ylab('') +
  theme_minimal() + labs(fill='mean corr')

```

## Plot network
### Get network back into matrix format
```{r}
fosnet.mat <-
  filter(fosnet,condition=="ALL") %>%
  select(seed,target,connection) %>%
  spread(target,connection)
row.names(fosnet.mat) <- fosnet.mat$seed
fosnet.mat <- select(fosnet.mat,-seed)
```

### Create the plot

```{r, message=FALSE, warning=FALSE}

thresh <- .2 # arbitrary, just for visualization
set.seed(5) # for reproducibility

plotNetwork <- function(mat,seedkey,threshold=thresh) {
    mat[mat<threshold] <- 0
    netplot <- network(mat)
    
    nodenames <- colnames(mat)
    nodegroups <- data.frame(seed = nodenames) %>%
      left_join(seedkey) %>%
      select(seed.module) 
    nodeoutline <- ifelse(nodenames=="AI"|nodenames=="DI"|nodenames=="GI","Insula","aa")
    
    netplot %v% "name" = nodenames
    netplot %v% "group" = nodegroups$seed.module
    netplot %v% "ins" = nodeoutline

    print(ggnet2(netplot,label="name",label.size=1,color = "group",shape="ins",
                 palette = "Accent", 
                 size = 0) +
            geom_point(aes(alpha = shape), size = 11) + scale_alpha_manual(values=c(0,0,1)) +
            geom_point(aes(color = color), size = 9) +
            geom_text(aes(label = label), color = "black") +
            guides(alpha=FALSE,shape=FALSE,color=FALSE))

}

net <- plotNetwork(fosnet.mat,seedkey)

ggsave('figs/network.eps',device="eps")


```

This plot is just another way of showing the connection matrix above. Since this format looks better with fewer connections, I applied an arbitrary threshold of `r thresh`.

# Plot node metrics, ranked by value
## Participation coefficient
```{r, fig.height=1.5, fig.width=3}

nodevals %>%
  filter(condition=='ALL') %>%
  mutate(seed = node) %>%
  left_join(seedkey) %>%
  filter(metric=="node_Ppos") %>%
  mutate(node.ordered=reorder(node,meanval),
         node.ordered=factor(node,levels=rev(levels(node.ordered)))) %>%
  mutate(seed.module = factor(seed.module)) %>%
  ggplot(aes(x=node.ordered,y=meanval)) + 
  geom_bar(stat="identity",aes(color=seed.ins),fill="gray70") +
  scale_color_manual(values=c("black","white")) +
  ylab('Participation coefficient') + 
  ggtitle('Participation coefficient - ALL') + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, color="black")) +
  xlab('') +
  guides(color=FALSE) -> p

print(p)

ggsave('figs/participation_coeff_ALL.eps',device="eps")

```


## Changes in participation coefficients
```{r, fig.height=4, fig.width=4}

nodevals %>%
  filter(metric=="node_Ppos") %>% 
  mutate(nodelab = factor(node,levels=rev(tmplevels))) %>%
  ggplot(aes(x=condition,y=meanval,fill=condition)) + geom_bar(stat="identity") +
  facet_wrap(~nodelab) + theme_minimal() + ylab('Participation coefficient') + xlab('')
  
```

# Average fos within module & relate to behavior

For this analysis, I'm going to summarize fos levels in the 3 insula regions. Then we can relate mean fos levels to behavior.

## Calculate average fos values
```{r, message=FALSE, warning=FALSE}

# summarize within modules & insula
rawdata.long %>%
  mutate(seed=region) %>%
  left_join(seedkey) %>%
  mutate(seed.module.ins=ifelse(seed.ins=="Insula",'insula',paste0('mod_',seed.module))) %>%
  filter(Condition=="ALL") %>%
  group_by(RAT,seed.module.ins) %>%
  summarize(meanFos = mean(fos,na.rm=TRUE)) %>%
  filter(!is.na(seed.module.ins))  -> modulemeans

# join with behavioral data
behavvars <-
  filter(rawdata,Condition=="ALL") %>%
  mutate(stress=ifelse(stress==1,'naive','stressed'),
         age=ifelse(age==1,'juvenile','adult')) %>%
  select(1:9) 

modulemeans <-
  left_join(modulemeans,behavvars)

```

## Show correlation with behavioral measures

```{r, fig.height=1.5, fig.width=2}

modulemeans %>%
  filter(seed.module.ins=="insula") %>%
  ggplot(aes(x=meanFos,y=Social.Exploration)) + 
  stat_smooth(method="lm",se=FALSE,color="gray") + 
  geom_point(aes(color=age,shape=stress),size=2) + 
  scale_color_manual(values=c("#bb996a","#97132c","#bb996a","#97132c")) +
  ggtitle('Social exploration by mean fos') + theme_minimal() +
  xlab('Insula: mean fos') +
  ylab('Social exploration')

ggsave('figs/scatterplot_insulafos_se.eps',device="eps")

```



## Regression predicting behavior

Now let's run some actual stats relating fos to behavior.

```{r}

modulemeans %>%
  group_by(seed.module.ins) %>%
  mutate(zmeanFos = (meanFos - mean(meanFos,na.rm=TRUE))/sd(meanFos,na.rm=TRUE)) %>%
  mutate(group = factor(interaction(stress,age)),
         condition = factor(ifelse(group=="naive.adult","Naive adult",
                                   ifelse(group=="naive.juvenile","Naive juvenile",
                                          ifelse(group=="stressed.adult","Stressed adult",
                                                 ifelse(group=="stressed.juvenile","Stressed juvenile",NA)))),
                            levels=c("Naive juvenile","Naive adult","Stressed juvenile","Stressed adult")),
         age = factor(age,levels=c("juvenile","adult"))) -> modulemeans


modulemeans %>%
  select(RAT,seed.module.ins,zmeanFos,Social.Exploration,stress,age) %>%
  spread(seed.module.ins,zmeanFos) -> statdata

model = lm(Social.Exploration ~ insula*stress*age, data=statdata)
anova(model)

```

